Flattering Advice: Avoiding Disappointment as a Driver of Gender Discrimination

### 2024-05-22 manuscript/manuscript.Rmd
Upload new Study 1

study1_math50 <- read.csv(here::here("Data/Cleaned Data/S1_Math50_cleaned.csv"))
study1_time1All <- readRDS(here::here("Data/Study1_time1All.Rds"))
study1_time1 <- readRDS(here::here("Data/Study1_time1.Rds"))
study1_time2_giver <- readRDS(here::here("Data/Study1_time2_giver.Rds"))
study1_time2_giver50 <- readRDS(here::here("Data/Study1_time2_giver50.Rds"))
study1_advice <- readRDS(here::here("Data/Study1_time2_advice.Rds"))
study1_time3 <- readRDS(here::here("Data/Study1_time3.Rds"))

@@

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study1_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study1_math50$Score[study1_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

@@

sortedScore <- sort(study1_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

@@

As part of our main experiment, participants choose to compete against a group
of previous participants and have their expectations anchored based on their performance.
Thus, we begin by first recruiting a sample of `r scales::comma(dim(study1_math50)[1])`
participants from Prolific to complete a 10-question multiple choice
mathematics quiz. The questions were taken from a paper-version of the
ASVAB standardized exam, such that answers were not available online.
Participants had five minutes to answer the quiz and were paid 10 cents for
each correctly answered question. We define the top 20 scorers as the
"High Performers" and the bottom 20 scorers as the "Low Performers." On
average, participants answered `r round(mean(study1_math50$Score),2)`
questions correctly, High Performers scored between `r HP_lower` and
`r HP_upper`, and Low Performers scored between `r LP_lower` and
`r LP_upper`. In order to anchor the expectations of participants in our main
experiment, we simulated 1,000 pairings of groups of five
participants, with the 5<sup>th</sup> percentile of groups scoring an average of
`r p5` and the 95 th percentile scoring an average of `r p95`. We report these
averages to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study1_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two
male participants to complete the survey, ending up with a sample of 500
men and 500 women ($M_{\text{Age}}$ = `r mean(study1_time1$Age)`). 
Participants completed the same 10-item mathematics quiz as the
earlier participants and were informed that their performance would
affect their bonus earnings in a follow-up stage to be conducted a few
days later. After completing the quiz, we informed them of the average
score of a group of five participants from the preliminary survey. We
randomly assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the 95<sup>th</sup>
percentile of groups, which scored `r p95` ("High Expectations"
treatment). Participants then made a guess (unincentivized) about how
many questions they think they answered correctly. The survey concluded
with basic demographic questions.

We then recruited `r scales::comma(dim(study1_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had to recommend
whether an advisee should compete against the Low Performers or the High
Performers (we used these terms in the survey). We anticipated that
being told to compete against High Performers is more flattering and hence being
told to compete against the Low Performers would be disappointing if one had
expected to do well.
Advisees would earn 30 cents if their score was equal to or higher than
a randomly selected member of the Low Performer group and they chose to
compete against this group, or 50 cents if they performed as well or
better than a randomly selected member of the High Performer group and they
chose to compete against that group. If
they scored lower, they would not receive any bonus.

@@

Finally, we invited participants from Stage 1
back for the follow-up survey. Following our preregistration, we kept
the survey open for 7 days. In total,
`r scales::comma(dim(study1_time3)[1])` participants
(`r scales::comma(dim(study1_time3[study1_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study1_time3[study1_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how many
questions they guessed they had answered correctly. Importantly, they
were not informed of their true score or the score of the groups they
could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

@@

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study1_time1[study1_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study1_time1[study1_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study1_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study1_time1[study1_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study1_time1[study1_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study1_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to our
expectations, we did find a gender difference in performance: men scored
`r round(mean(study1_time1[study1_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study1_time1[study1_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study1_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more questions 
correctly than did
women
(`r round(mean(study1_time1[study1_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study1_time1[study1_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study1_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We observe
that women underestimate their performance by 0.76 points, while men do so by
only 0.27 points `r papaja::apa_print(t.test(overconfidence ~ Gender, data = study1_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more confident in
their performance than are women, as we had expected.

```{r study1regs}
study1_advice <- study1_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study1_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study1_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study1_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.")

@@

ProbHG <- study1_advice %>%
  group_by(Condition, Performance, Gender) %>%
  dplyr::summarize(avg_hardGroup = mean(Advice))

@@

study1_time3 <- study1_time3 %>% dplyr::mutate(AdviseeCondition = factor(AdviseeCondition, levels = c('Low', 'High'))) %>%
     dplyr::mutate(Adopt = ifelse(Bet == Advice,1,0))

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study1_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study1_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study1_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study1_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect.")

>>>>

study3_math50 <- read.csv(here::here("Data/Cleaned Data/S3_Math50_cleaned.csv"))
study3_time1All <- readRDS(here::here("Data/Study3_time1All.Rds"))
study3_time1 <- readRDS(here::here("Data/Study3_time1.Rds"))
study3_time2_giver <- readRDS(here::here("Data/Study3_time2_giver.Rds"))
study3_time2_giver50 <- readRDS(here::here("Data/Study3_time2_giver50.Rds"))
study3_advice <- readRDS(here::here("Data/Study3_time2_advice.Rds"))
study3_time3 <- readRDS(here::here("Data/Study3_time3.Rds"))

@@

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study3_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study3_math50$Score[study3_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

@@

sortedScore <- sort(study3_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

@@

As part of our main experiment, participants choose to compete against a group
of previous participants and have their expectations anchored based on their performance.
Thus, we begin by first recruiting a sample of `r scales::comma(dim(study3_math50)[1])`
participants from Prolific to complete a 10-question multiple choice
mathematics quiz. The questions were taken from a paper-version of the
ASVAB standardized exam, such that answers were not available online.
Participants had five minutes to answer the quiz and were paid 10 cents for
each correctly answered question. We define the top 20 scorers as the
"High Performers" and the bottom 20 scorers as the "Low Performers." On
average, participants answered `r round(mean(study3_math50$Score),2)`
questions correctly, High Performers scored between `r HP_lower` and
`r HP_upper`, and Low Performers scored between `r LP_lower` and
`r LP_upper`. In order to anchor the expectations of participants in our main
experiment, we simulated 1,000 pairings of groups of five
participants, with the 5<sup>th</sup> percentile of groups scoring an average of
`r p5` and the 95 th percentile scoring an average of `r p95`. We report these
averages to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study3_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two
male participants to complete the survey, ending up with a sample of 500
men and 500 women ($M_{\text{Age}}$ = `r mean(study3_time1$Age)`). 
Participants completed the same 10-item mathematics quiz as the
earlier participants and were informed that their performance would
affect their bonus earnings in a follow-up stage to be conducted a few
days later. After completing the quiz, we informed them of the average
score of a group of five participants from the preliminary survey. We
randomly assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the 95<sup>th</sup>
percentile of groups, which scored `r p95` ("High Expectations"
treatment). Participants then made a guess (unincentivized) about how
many questions they think they answered correctly. The survey concluded
with basic demographic questions.

We then recruited `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had to recommend
whether an advisee should compete against the Low Performers or the High
Performers (we used these terms in the survey). We anticipated that
being told to compete against High Performers is more flattering and hence being
told to compete against the Low Performers would be disappointing if one had
expected to do well.
Advisees would earn 30 cents if their score was equal to or higher than
a randomly selected member of the Low Performer group and they chose to
compete against this group, or 50 cents if they performed as well or
better than a randomly selected member of the High Performer group and they
chose to compete against that group. If
they scored lower, they would not receive any bonus.

@@

Finally, we invited participants from Stage 1
back for the follow-up survey. Following our preregistration, we kept
the survey open for 7 days. In total,
`r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how many
questions they guessed they had answered correctly. Importantly, they
were not informed of their true score or the score of the groups they
could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

@@

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to our
expectations, we did find a gender difference in performance: men scored
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more questions 
correctly than did
women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We observe
that women underestimate their performance by 0.76 points, while men do so by
only 0.27 points `r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more confident in
their performance than are women, as we had expected.

```{r study1regs}
study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.")

@@

ProbHG <- study3_advice %>%
  group_by(Condition, Performance, Gender) %>%
  dplyr::summarize(avg_hardGroup = mean(Advice))

@@

study3_time3 <- study3_time3 %>% dplyr::mutate(AdviseeCondition = factor(AdviseeCondition, levels = c('Low', 'High'))) %>%
     dplyr::mutate(Adopt = ifelse(Bet == Advice,1,0))

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect.")
